Early Observations of \abc:\\ \textbf{a Type Ia Supernova Featuring Either} Strong Ejecta Mixing and/or Interaction with Diffuse Material

### 2017-10-26 doc/ms.tex
Reword abstract and include estimate of M_Ni from L_max

\title{Early Observations of the Type Ia Supernova \abc:\\
Evidence for Strong Ejecta Mixing and/or Interaction with Diffuse Material}

@@

Early observations of Type Ia supernovae (SNe) provide a unique probe of their progenitor systems and explosion physics. Here, we report the intermediate Palomar Transient Factory (iPTF) discovery of an extraordinarily young SN Ia, \abc. By fitting a power law to our early light curve, we infer that first light for the SN, that is when appreciable heating from $^{56}$Ni decay has first diffused to the photosphere, only occurred $0.15\pm_{0.07}^{0.15}$~d before our first detection. In the $\sim$24 hr after discovery, \abc\ rose by $\sim$2 mag, following a near-linear rise in flux for $\gtrsim$3 d. 
Strong \ion{C}{2} absorption is detected in the early spectra of \abc, before disappearing after $\sim$7~d. Unlike the extensively-observed Type Ia SN\,2011fe, the $(B-V)_0$ colors of \abc\ are blue and nearly constant in the days after explosion. We show that our early observations of \abc\ cannot be explained by either SN shock breakout and the associated, subsequent cooling, or the SN ejecta colliding with a stellar companion. Instead, we argue that the early characteristics of \abc, including: (i) the rapid, near-linear rise, (ii) the non-evolving blue colors, and (iii) the strong absorption from ionized carbon, are the result of either vigorous mixing of radioactive $^{56}$Ni in the SN ejecta, or ejecta interaction with diffuse material, or a combination of the two. In the next few years, dozens of very young \textit{normal} SNe Ia will be discovered, and observations similar to those presented here will constrain the white dwarf explosion mechanism.

@@

\begin{figure*}[!htb]
  \centering
  \includegraphics[width=0.78\textwidth]{spectra.pdf}
  \caption{Observed spectral sequence of \abc. The spectra are 
  normalized by their median flux between 6,000 and 
  7,000$\,\textrm{\AA}$.  The phase of each spectrum 
  relative to $T_{B_\mathrm{max}}$ is shown. Telluric absorption
  bands are grayed out. Line identifications are provided for 
  the spectral features discussed in the text. 
  For clarity, high-resolution spectra obtained with the VLT have been omitted (see Ferretti et al.\ 2017, submitted, for a detailed discussion of these spectra).}
  \label{fig:spec_seq}
\end{figure*}

@@

To determine the brightness and time of $B$-band maximum for 
\abc, we fit the P60 light curves with the \texttt{sncosmo} software package.\footnote{\texttt{sncosmo} is available at
  \url{https://sncosmo.readthedocs.io}.} This fit includes a \texttt{SALT2} template \citep{2007A&A...466...11G} that has been corrected for extinction using the \citep{1999PASP..111...63F} reddening law, $R_V=3.1$, and $E(B-V) = 0.08 \; \mathrm{mag}$.

@@

We measure the (pseudo)bolometric luminosity, $L_{UVOIR}$, of \abc\ at peak via trapezoidal integration of the reddening-corrected flux from the UV, optical, and NIR (UVOIR) filters. The light curves in the individual filters are interpolated so that $L_{UVOIR}$ is evaluated at common epochs in each filter. From this integration, we determine that \abc\ reaches a maximum luminosity $L_\mathrm{max} = 1.2 \pm 0.1 \times 10^{43} \; \mathrm{erg \; s}^{-1}$. This value is consistent with the normal SNe Ia studied in  \citet{2016A&A...588A..84D}, a sample which excludes low-luminosity SNe.

@@

Multiband observations of \abc\ began $\sim$1.5 d after discovery, which allows us to trace its color evolution starting $\sim$1.7 d after $t_0$. In Figure~\ref{fig:B-Vcolors} we compare the $(B - V)_0$ color evolution of \abc\ and SN\,2011fe, where the observations of SN\,2011fe are taken from \citet{2016ApJ...820...67Z}. For both SNe the colors have been corrected for the total inferred reddening along the line of sight. Interestingly, \abc\ has a nearly flat color evolution up to 10 d before $T_{B,\mathrm{max}}$, while SN\,2011fe exhibits red colors initially before evolving to the blue.

>>>>

\title{Early Observations of \abc:\\
\textbf{a Type Ia Supernova Featuring Either} Strong Ejecta Mixing and/or Interaction with Diffuse Material}

@@

Early observations of Type Ia supernovae (SNe) provide a unique probe of their
progenitor systems and explosion physics. Here, we report the intermediate
Palomar Transient Factory (iPTF) discovery of an extraordinarily young SN Ia,
\abc. By fitting a power law to our early light curve, we infer that first
light for the SN, \textbf{that is when the SN produces appreciable optical
radiation following shock breakout}, \textbf{occurred only}
$0.15\pm_{0.07}^{0.15}$\,d before our first detection. In the $\sim$24\,hr
after discovery, \abc\ rose by $\sim$2\,mag, \textbf{featuring} a near-linear
rise in flux for $\gtrsim$3 d. \textbf{Early spectra show strong \ion{C}{2}
absorption, which disappears after $\sim$7\,d.} Unlike the
extensively-observed Type Ia SN\,2011fe, the $(B-V)_0$ colors of \abc\ are
blue and nearly constant in the days after explosion. We show that our early
observations of \abc\ cannot be explained by either SN shock breakout and the
associated, subsequent cooling, or the SN ejecta colliding with a stellar
companion. Instead, we argue that the early characteristics of \abc,
including: (i) the rapid, near-linear rise, (ii) the non-evolving blue colors,
and (iii) the strong absorption from ionized carbon, are the result of either
vigorous mixing of radioactive $^{56}$Ni in the SN ejecta, or ejecta
interaction with diffuse material, or a combination of the two. In the next
few years, dozens of very young \textit{normal} SNe Ia will be discovered, and
observations similar to those presented here will constrain the white dwarf
explosion mechanism.

@@

\begin{figure*}[!htb]
  \centering
  \includegraphics[width=0.78\textwidth]{spectra.pdf}
  \caption{Observed spectral sequence of \abc. The spectra are 
  normalized by their median flux between 6,000 and 
  7,000$\,\textrm{\AA}$.  The phase of each spectrum 
  relative to \textbf{the time of $B$-band maximum} is shown. Telluric absorption
  bands are grayed out. Line identifications are provided for 
  the spectral features discussed in the text. 
  For clarity, high-resolution spectra obtained with the VLT have been omitted (see Ferretti et al.\ 2017, submitted, for a detailed discussion of these spectra).}
  \label{fig:spec_seq}
\end{figure*}

@@

To determine the brightness and time of $B$-band maximum for 
\abc, we fit the P60 light curves with the \texttt{sncosmo} software package.\footnote{\texttt{sncosmo} is available at
  \url{https://sncosmo.readthedocs.io}.} This fit includes a \texttt{SALT2} template \citep{2007A&A...466...11G} that has been corrected for extinction using the \citet{1999PASP..111...63F} reddening law, $R_V=3.1$, and $E(B-V) = 0.08 \; \mathrm{mag}$.

@@

We measure the (pseudo)bolometric luminosity, $L_{UVOIR}$, of \abc\ at peak via trapezoidal integration of the reddening-corrected flux from the UV, optical, and NIR (UVOIR) filters. The light curves in the individual filters are interpolated so that $L_{UVOIR}$ is evaluated at common epochs in each filter. From this integration, we determine that \abc\ reaches a maximum luminosity $L_\mathrm{max} = 1.2 \pm 0.1 \times 10^{43} \; \mathrm{erg \; s}^{-1}$. This value is consistent with the normal SNe Ia studied in  \citet{2016A&A...588A..84D}, a sample which excludes low-luminosity SNe. \textbf{Following Arnett's rule \citet{1982ApJ...253..785A, 1985Natur.314..337A}, it is possible to derive the mass of $^{56}$Ni synthesized in the explosion from $L_\mathrm{max}$. Assuming a rise time of $19 \pm 3$\,d (see \citealt{2006A&A...450..241S}) we find $M_\mathrm{Ni} = 0.6 \pm 0.1 M_\odot$.\footnote{Assuming a rise time of 17.9\,d, see \S\ref{sec:lc_fit} results in a consistent estimate of $M_\mathrm{Ni}$.}}

@@

Multiband observations of \abc\ began $\sim$1.5 d after discovery, which allows us to trace its color evolution starting $\sim$1.7 d after $t_0$. In Figure~\ref{fig:B-Vcolors} we compare the $(B - V)_0$ color evolution of \abc\ and SN\,2011fe, where the observations of SN\,2011fe are taken from \citet{2016ApJ...820...67Z}. For both SNe the colors have been corrected for the total inferred reddening along the line of sight. Interestingly, \abc\ has a nearly flat color evolution up to 10\,d before $T_{B,\mathrm{max}}$, while SN\,2011fe exhibits red colors initially before evolving to the blue.
